---
title: "Data Analytics Final Project"
output:
  pdf_document: default
  html_notebook: default
---
## Global City Population Estimates

### Introduction
* Description: of Urban Agglomerations with 300,000 Inhabitants or More 1950-2030 (thousands)

* Publisher: United Nations, Department of Economic and Social Affairs, Population Division (2014). World Urbanization Prospects: The 2014 Revision

* Source URL: https://data.gov.uk/dataset/128a0e7c-e51e-4fba-b743-289e2a8debdf/global-city-population-estimates

* Population of 1692 cities in 1950 – 2010 (temporal: 5 yrs)

* Estimate for 2015 - 2030

### Goals
* History trend, future estimate for each city, continent 

* Population growth animation (top 20 cities each decade) + prediction of next several decades

* Population clustering in each continent, or one particular country (map)

* Cities/countries with the most/least increasing rate 

* Types of population growth  

```{r message=FALSE, warning=FALSE}
# libraries 
library(readxl)
library(dplyr)
library(plyr)
suppressMessages(library(dplyr))
library(scales)
library(ggplot2)
library(rworldmap)
library(nlme)
library(minpack.lm)
library(car)
library(growthrates)
library(growthcurver)
library(tidyverse)
library(class)
library(rpart)
library(rpart.plot)
library(maps)
```

### 1. Read data and basic information
```{r}
# read
Pop_data_all = read_excel("/Users/wangfeng/Desktop/ITWS 6600 Data Analytics/Project/global-city-population-estimates.xls", sheet = 2)

# Number of countries
num_country = nrow(as.data.frame(table(Pop_data_all$`Country or area`)))
cat('There are', num_country, 'countries in this dataset.\n')

# Number of cities
num_city = nrow(Pop_data_all)
cat('There are', num_city, 'cities in this dataset.\n')
cat('The dataset has history data from 1950 - 2010 and estimates for 2015 - 2030 with a temporal of 5 years.')

```

### 2. Exploratory Data Analysis
```{r}
# Plot all the cities
worldmap = getMap(resolution = "li")
plot(worldmap)
points(Pop_data_all$Longitude, Pop_data_all$Latitude, col = "lightblue", pch = 20, cex = 0.6)

k.max = 12
wss = sapply(1:k.max,function(k){kmeans(Pop_data_all[,6:7], k, nstart = 20, iter.max = 20)$tot.withinss})
plot(1:k.max,wss, type= "b", xlab = "Number of clusters(k)", ylab = "Within cluster sum of squares")
icluster = kmeans(Pop_data_all[,6:7], 6, nstart = 7, iter.max = 20)

# plot 
worldmap = getMap(resolution = "li")
plot(worldmap)
points(Pop_data_all$Longitude, Pop_data_all$Latitude, col = "lightblue", pch = 20, cex = 0.6)
centers = as.data.frame(icluster$centers)
points(centers$Longitude, centers$Latitude, col = "brown1", pch = 20, cex = 2.5)
```


```{r}
# Plot all the cities with color indicating population in 2010
worldmap = getMap(resolution = "li")
plot(worldmap)
firstgroup = Pop_data_all[which(Pop_data_all$`2010` > 10000),]
secondgroup = Pop_data_all[which(Pop_data_all$`2010` > 5000 | Pop_data_all$`2010` < 10000),]
thirdgroup = Pop_data_all[which(Pop_data_all$`2010` < 5000),]
points(firstgroup$Longitude, firstgroup$Latitude, col = "brown1", pch = 20, cex = 1.5)
#points(secondgroup$Longitude, secondgroup$Latitude, col = "royalblue1", pch = 20, cex = 0.4)
#points(thirdgroup$Longitude, thirdgroup$Latitude, col = "lightgreen", pch = 20, cex = 0.2)

```

```{r include=FALSE}
# Top 500 cities in each temporal
worldmap2 = getMap(resolution = "li")
#for(i in 8:20){
#  top200_list = Pop_data_all[order(-Pop_data_all[[i]]),][1:200,]
#  plot(worldmap2)
#  points(top200_list$Longitude, top200_list$Latitude, col = "lightblue", pch = 20, cex = 0.7)
#}
```


```{r}
# Countries 
country_table = as.data.frame(table(as.factor(Pop_data_all$`Country or area`)))
country_table = filter(country_table, Freq > 20)
country_table = country_table[order(country_table$Freq),]
# barplot(country_table$Freq, 
#        names.arg = country_table$Var1, main = 'Countries',
#        col = 'lightblue',
#        horiz = TRUE, cex.names = 0.5, las = 1,
#        xlim = c(0,400))
#axis(1, at = seq(0, 400, 25))

ggplot(data = country_table, aes(x = Var1, y = Freq)) +
  geom_bar(stat = 'identity', fill="lightblue")+
  geom_text(aes(label = Freq), vjust=1, color="black", size=3)+
  ylab("Frequency")+
  xlab("Country") +
  ggtitle(paste("Most resided countries"))+
  scale_x_discrete(limits = country_table$Var1)+
  theme_minimal() + 
  coord_flip()
```

#### 2.1 Twenty most resided cities every five years
Using prediction model to interpolate the values in between and make it a smooth animation! 
```{r warning=FALSE}
# Select the most resided cities and draw barplots
for (i in c(8,20)){
  year = colnames(Pop_data_all)[i]
  temp_data = Pop_data_all
  temp_data = temp_data[order(-temp_data[[year]]),] # Using [[]] to call columns with a variable
  temp_20 = temp_data[1:20,]
  temp_20 = temp_20[order(temp_20[[year]]),] # For barplot in descending order
  # barplot(temp_20[[i]], 
  #         main=paste("Population of top 20 resided cities in", year), 
  #         names.arg = temp_20$`Urban Agglomeration`, 
  #         horiz = TRUE, cex.names = 0.5, las = 2, col = 'lightblue',
  #         cex.axis = 0.8,
  #         xlab = 'Popolation (in thousands)',
  #         xlim = c(0, 38000)) 
  # paste to include variable
  # cex.names to shrink the size 
  # las = 2 to rotate
  plot = ggplot(data = temp_20, aes(x = `Urban Agglomeration`, y = temp_20[[i]])) +
  geom_bar(stat = 'identity', fill="lightblue")+
  geom_text(aes(label = temp_20[[i]]), vjust=1, color="black", size=3)+
  ylab("Population (in thousands)")+
  ylim(c(0,38000)) +
  ggtitle(paste("Top 20 resided cities in ", year))+
  scale_x_discrete(limits = temp_20$`Urban Agglomeration`)+
  theme_minimal() + 
  coord_flip()
  
  print(plot)
}
```

#### 2.2 Increasing rate
Finding the increasing rates with respect to five years ago and finding the 20 most and least cities. Please note that not all rates in every temporal are shown for brevity. 
```{r}
rise_rate = vector()
for (i in 9:20){
  year_new = colnames(Pop_data_all)[i]
  year_old = colnames(Pop_data_all)[i-1]
  
  rise_table = Pop_data_all[,4]
  rise_table = mutate(rise_table, increasing = 
                      (Pop_data_all[[year_new]] - Pop_data_all[[year_old]])/
                                (Pop_data_all[[year_old]] + 1)) # plus one to prevent infinity  
  rise_rate[i-8] = mean(unlist(rise_table$increasing))
  
  # Increasing rate - highest 20
  rise_table_high = rise_table[order(-rise_table$increasing),][1:20,]
  rise_table_high$increasing = rise_table_high$increasing * 100
#   barplot(rise_table_high$increasing, 
#          main = paste("Top 20 cities with the highest increasing rate from", year_old, "to",
#                      year_new), 
#          names.arg = rise_table_high$`Urban Agglomeration`, 
#          horiz = TRUE, cex.names = 0.5, las = 2, col = 'lightblue',
#          cex.axis = 0.8,
#          xlab = 'Increasing rate (percentage)')
  
  # Increasing rate - lowest 20
  rise_table_low = rise_table[order(rise_table$increasing),][1:20,]
  rise_table_low$increasing = rise_table_low$increasing * 100
 # barplot(rise_table_low$increasing, 
 #         main = paste("Top 20 cities with the lowest increasing rate from", year_old, "to",
#                      year_new), 
#          names.arg = rise_table_low$`Urban Agglomeration`, 
#          horiz = TRUE, cex.names = 0.5, las = 2, col = 'lightblue',
#          cex.axis = 0.8,
#          xlab = 'Increasing rate (percentage)')
}
```

```{r}
# World average increasing rate 
plot(colnames(Pop_data_all)[9:20], rise_rate, type = 'o', col = 'steelblue', lwd = 3,
     main = 'World average increasing rate ',
     xla = 'Year', ylab = 'Increasing rate')
```

#### 2.4 History trend
To get a first sight of the history data, the trends of the 20 most resided cities in 2010 in the past half a century are investigated. 
```{r}
# Top 20 resided
Pop_data_2010 = Pop_data_all[order(-Pop_data_all$`2010`),][1:200,]
for (i in 1:3){
  city = Pop_data_2010$`Urban Agglomeration`[i]
  temp_data = t(Pop_data_2010[i, 8:20])
  plot(rownames(temp_data), temp_data, 
       type = "o", 
       main = paste("Historical increasing trend from 1950 to 2010 in", city),
       col = 'steelblue', lwd = 3,
       xlab = 'Year', ylab = 'Population')
}
```

```{r}
# Three stages in one plot
# Shanghai, Mexico city, Osaka
Pop_data_2010 = Pop_data_all[order(-Pop_data_all$`2010`),][,1:20]
stage_3 = data.frame(Time = colnames(Pop_data_2010)[8:20], Shanghai = t(Pop_data_2010[4,8:20]),
                     MexicoCity = t(Pop_data_2010[3,8:20]), Osaka = t(Pop_data_2010[6,8:20]))
colors = c('Shanghai' = "lightblue", 'Mexico City' = "brown1", 'Osaka' = "steelblue")
ggplot(stage_3, aes(x = Time)) + 
  geom_line(aes(y = Shanghai, group = 1, color = 'Shanghai'), lwd = 1.5) +
  geom_point(aes(y = Shanghai, group = 1), color = "white") +
  geom_line(aes(y = MexicoCity, group = 1, color = 'Mexico City'), lwd = 1.5) +
  geom_point(aes(y = MexicoCity, group = 1), color = "white") +
  geom_line(aes(y = Osaka, group = 1, color = 'Osaka'), lwd = 1.5) +
  geom_point(aes(y = Osaka, group = 1), color = "white") +
  xlab('Year') +
  ylab('Population (in thousands)') +
  ggtitle('Population growth in Shanghai, Mexico City and Osaka') +
  scale_color_manual(name = 'Cities', values = colors)
  
```

```{r}
# Least 20 resided
Pop_data_2010 = Pop_data_all[order(Pop_data_all$`2010`),][,1:20]
for (i in 1:20){
  city = Pop_data_2010$`Urban Agglomeration`[i]
  temp_data = t(Pop_data_2010[i, 8:20])
#  plot(rownames(temp_data), temp_data, 
#       type = "o", 
#       main = paste("Population history trend from 1950 to 2010 in", city),
#       col = 'steelblue', lwd = 3,
#       xlab = 'Year', ylab = 'Population')
}
```
#### 2.5 World population analysis
```{r include=FALSE}
# The histogram of all cities in different years
for(i in seq(8, 20)){
  year = colnames(Pop_data_all)[i]
#  hist(unlist(Pop_data_all[,i]), main = paste('Histogram of all the cities in ', year),
#       breaks = 100, xlim = c(0, 35000), ylim = c(0, 1300))
}
```

```{r}
# Mean population of all the cities in the past 60 years
mean_vec = vector()
min_vec = vector()
max_vec = vector()
for(i in seq(1,13)){
  mean_vec[i] = mean(unlist(Pop_data_all[,7+i]))
  min_vec[i] = min(Pop_data_all[,7+i])
  max_vec[i] = max(Pop_data_all[,7+i])
}
stat_allcities = data.frame(Year = as.numeric(colnames(Pop_data_all)[8:20]), 
                            Mean = mean_vec, Min = min_vec, Max = max_vec)
par(mfrow = c(1,3))
plot(stat_allcities$Year, stat_allcities$Max, type = 'o', col = 'steelblue', main = 'Max', lwd = 2)
plot(stat_allcities$Year, stat_allcities$Mean, type = 'o', col = 'steelblue', main = 'Mean', lwd = 2)
plot(stat_allcities$Year, stat_allcities$Min, type = 'o', col = 'steelblue', main = 'Min', lwd = 2)


```




### 3. Population Growth Model
```{r}
model_vec = rep(0, nrow(Pop_data_all)) 
Pop_model = data.frame(Pop_data_all[c(2,4,6:20)])
Pop_model$Model = model_vec
# Logistic: 1
# Exponential: 0
```

There are three kinds of growth model for reference, linear model, exponential model and logistic model. With the results above, it can be seen that in the beginning of a development of a city, it is usually exponential. And it will be logistic when a city is expanding to a reletively large scale. The linear growth model is not very common hence is not used here. 

The next problem is how to identify which model the population growth follows. 
#### 3.1 Logistic or Exponential?
In this section, the model of the city population growth is determined and the predictions are made. Here logistic model is realized by growthcurver model. 
```{r}
start_time <- Sys.time()

num_cities = c(1:166, 168:269, 271:368, 370:480, 482:920, 922:945, 947:1051, 1053:1338, 1340:1493, 1495:1692)
for(i in num_cities) # nrow(Pop_data_all))
{
# Start
city_name = Pop_data_all$`Urban Agglomeration`[i]
pop_time = as.data.frame(t(Pop_data_all[i,8:20]))
pop_time$Time = as.numeric(row.names(pop_time))
names(pop_time)[1] = 'Population'
pop_time = pop_time[,c(2,1)]

# Exponential model
exponential.model = lm(log(Population) ~ Time, data = pop_time)
exp_pred = exp(predict(exponential.model, data.frame(Time = pop_time$Time)))
exp_SSE = sum((exp_pred - pop_time$Population)**2)

# Logistic model 
# logistic.model = nls(Population ~ (p0*K)/(p0 + (K - p0)*exp(-r*Time)), data = pop_time)
# Try to find some initial guesses
# init = coef(lm(log(Population) ~ Time, data = pop_time))
# init_list = list(phi1 = pop_time$Population[1], phi2 = as.numeric(init[1]), phi3 = as.numeric(init[2]))

# SSLogis : y = phi1/[1 + exp(-(t-phi2)/phi3)]
#control = nls.control(minFactor = 1/2048, printEval = TRUE)
#logistic.model = gnls(Population ~ SSlogis(Time, phi1, phi2, phi3), data = pop_time, control = control) #  control = nls.control(minFactor = 1/22),  start = init_list
#alpha = coef(logistic.model)
#logistic_pred = alpha[1]/(1 + exp(-(pop_time$Time - alpha[2])/alpha[3]))
# logistic_pred = (alpha[1]*alpha[2])/(alpha[1] + (alpha[2] - alpha[1])*exp(-alpha[3]*Time))
#logistic_SSE = sum((logistic_pred - pop_time$Population)**2)

# Using self made function
# Carrying capacity model: N =[N0 * exp(r*t)]/[1 + N0 * (exp(r*t) - 1)/K] 
# K : carrying capacity
# dN/dt = r*N*(1 - N/K)
#fit.logistic = function(par, data){
#  r = par[1]
#  K = par[2]
#  N = data[,2]
#  t = data[,1] - 1950
#  N0 = data[2,1]
#  temp = N0 *exp(r*t)/(1 + N0 * (exp(r*t)-1)/K)
#  sumsq = sum((N -temp)^2)
#}
#r_init = (pop_time[13,2] - pop_time[1,2])/(60 * pop_time[1,2]) # 60 yrs
#K_init = pop_time[13,2]
#N0_init = pop_time[1,2]
#par = c(N0_init, r_init, K_init)
#logistic.model = optim(par, fit.logistic, data = pop_time)
#r = logistic.model$par[2]
#K = logistic.model$par[3]
#N0 = logistic.model$par[1]
#logistic_pred = (N0 * exp(r*(pop_time$Time- 1950)))/(1 + N0 * (exp(r*(pop_time$Time- 1950)) - 1)/K)

# Growthcurver model
# K / (1 + ((K - N0) / N0) * exp(-r * t))
logistic.model = SummarizeGrowth(pop_time$Time-1950, pop_time$Population)
a = logistic.model$vals
logistic_pred = a$k / (1 + ((a$k - a$n0) / a$n0) * exp(- a$r * (pop_time$Time-1950))) + pop_time$Population[1]
logistic_SSE = sum((logistic_pred - pop_time$Population)**2)

# Compare the errors
a_mod = 'exponential model'
b_mod = 'logistic model'
new_time = data.frame(Time = c(2015, 2020, 2025, 2030))
all_time = c(pop_time$Time, new_time$Time)
if(exp_SSE > logistic_SSE){
  # logistic model
  model_vec[i] = 1
  Pop_model$Model[i] = 1
  # print(paste('The population in', i, city_name, 'is best to be modeled by', b_mod))
  # predict
  Pop_pred = a$k / (1 + ((a$k - a$n0) / a$n0) * exp(- a$r * (new_time$Time-1950))) + pop_time$Population[1]
  Pop_model[i,19:22] = Pop_pred
  all_pop = a$k / (1 + ((a$k - a$n0) / a$n0) * exp(- a$r * (all_time - 1950))) + pop_time$Population[1]
} else {
  # expoential model
  # print(paste('The population in', i, city_name, 'is best to be modeled by', a_mod))
  Pop_pred = exp(predict(exponential.model, newdata = data.frame(Time = c(2015, 2020, 2025, 2030))))
  Pop_model[i,19:22] = Pop_pred
  all_pop = c(exp_pred, Pop_pred)
}

# plot(pop_time$Time, pop_time$Population, pch = 16, main = paste('Population modeling in', city_name))
# lines(pop_time$Time, exp_pred, col = 'brown1', lwd = 2) # The results of prediction should be exp()
# lines(pop_time$Time, logistic_pred, col = "steelblue", lwd = 2)


# plot(pop_time$Time, pop_time$Population, pch = 16, main = paste('Population modeling in', city_name), 
#     xlim = c(1950,2030), ylim = c(min(all_pop), max(Pop_data_all[i,8:24])*1.1), lwd = 2)
# lines(all_time, all_pop, col = "brown1", lwd = 2)
# points(new_time$Time, Pop_data_all[i,21:24], col = 'steelblue')

# End 
}

# 9 of them cannot use exp or logistic model 
# Spline using loess is used
for(i in c(167, 270, 369, 481, 921, 946, 1052, 1339, 1494)) # nrow(Pop_data_all))
{
# Start
city_name = Pop_data_all$`Urban Agglomeration`[i]
pop_time = as.data.frame(t(Pop_data_all[i,8:20]))
pop_time$Time = as.numeric(row.names(pop_time))
names(pop_time)[1] = 'Population'
pop_time = pop_time[,c(2,1)]

# Spline model 
loess.model = loess(Population ~ Time, pop_time, control=loess.control(surface="direct")) # To extrapolate
loess_pred = predict(loess.model, data.frame(Time = c(2015, 2020, 2025, 2030)), se = TRUE)
Pop_model[i,19:22] = loess_pred$fit
all_time = c(pop_time$Time, new_time$Time)
all_pop = c(pop_time$Population, loess_pred$fit)

# plot(pop_time$Time, pop_time$Population, pch = 16, main = paste('Population modeling in', city_name), 
#      xlim = c(1950,2030), ylim = c(min(all_pop), max(Pop_data_all[i,8:24])*1.1), lwd = 2)
# lines(all_time, all_pop, col = "brown1", lwd = 2)
# points(new_time$Time, Pop_data_all[i,21:24], col = 'steelblue')

model_vec[i] = 2
Pop_model$Model[i] = 2
# End 
}

names(Pop_model)[19] = 'X2015'
names(Pop_model)[20] = 'X2020'
names(Pop_model)[21] = 'X2025'
names(Pop_model)[22] = 'X2030'


end_time <- Sys.time()
end_time - start_time
```

```{r}
# Models
table(Pop_model$Model)
# 0 : exponential
# 1 : logistic
# 2 : spline
```

Most of them follow logistic model but some of them are better to use exponetial model to predict. 

#### 3.2 Growthcurver or SSlogis
In this section, SSlogis model is used. But many of the cites cannot use this model. The plan is to use two kinds of the logistic model and to choose the more accurate one. But too many cities have difficulties using SSlogis. So this model is abandoned. 
```{r}
Pop_model2 = Pop_model
# Logistic model
start_time <- Sys.time()
for(i in seq(1,1)) # nrow(Pop_data_all))
{
print(i)
# Start
city_name = Pop_data_all$`Urban Agglomeration`[i]
pop_time = as.data.frame(t(Pop_data_all[i,8:20]))
pop_time$Time = as.numeric(row.names(pop_time))
names(pop_time)[1] = 'Population'
pop_time = pop_time[,c(2,1)]
# SSLogis : y = phi1/[1 + exp(-(t-phi2)/phi3)]

logistic.model = nls(Population ~ SSlogis(Time, phi1, phi2, phi3), data = pop_time)
a = as.numeric(coef(logistic.model))
new_time = data.frame(Time = c(2015, 2020, 2025, 2030))
Pop_pred = SSlogis(new_time$Time, a[1], a[2], a[3])
Pop_model2[i,19:22] = Pop_pred
all_time = c(pop_time$Time, new_time$Time)
all_pop = SSlogis(all_time, a[1], a[2], a[3])

#plot(pop_time$Time, pop_time$Population, pch = 16, main = paste('Population modeling in', city_name), 
#     xlim = c(1950,2030), ylim = c(min(all_pop), max(Pop_data_all[i,8:24])*1.1))
#lines(all_time, all_pop, col = "red")
#points(new_time$Time, Pop_data_all[i,21:24], col = 'green')
# End 
}
end_time <- Sys.time()
end_time - start_time

```

```{r}
# Using spline model LOESS
Pop_model3 = Pop_model
start_time <- Sys.time()
for(i in seq(1,1)) # nrow(Pop_data_all))
{
# print(i)
# Start
city_name = Pop_data_all$`Urban Agglomeration`[i]
pop_time = as.data.frame(t(Pop_data_all[i,8:20]))
pop_time$Time = as.numeric(row.names(pop_time))
names(pop_time)[1] = 'Population'
pop_time = pop_time[,c(2,1)]
new_time = data.frame(Time = c(2015, 2020, 2025, 2030))

loess.model = loess(Population ~ Time, pop_time, control=loess.control(surface="direct")) # To extrapolate
loess_pred = predict(loess.model, data.frame(Time = c(2015, 2020, 2025, 2030)), se = TRUE)
Pop_model3[i,19:22] = loess_pred$fit
all_time = c(pop_time$Time, new_time$Time)
all_pop = c(pop_time$Population, loess_pred$fit)

#plot(pop_time$Time, pop_time$Population, pch = 16, main = paste('Population modeling in', city_name), 
#     xlim = c(1950,2030), ylim = c(min(all_pop), max(Pop_data_all[i,8:24])*1.1))
#lines(all_time, all_pop, col = "red")
#points(new_time$Time, Pop_data_all[i,21:24], col = 'green')
# End 
}
names(Pop_model3)[19] = 'X2015'
names(Pop_model3)[20] = 'X2020'
names(Pop_model3)[21] = 'X2025'
names(Pop_model3)[22] = 'X2030'

end_time <- Sys.time()
end_time - start_time
```

#### 3.3 Prediction errors
In this section, the predictions are compared with the given estimates. The estimates in the dataset are supposed to be predicted combining many factors instead of just data analysis.
```{r}
# Using logistic model
# Error compared with the given estimates
Error2015 = abs(Pop_model$X2015 - Pop_data_all$`2015`)/Pop_data_all$`2015`
Error2020 = abs(Pop_model$X2020 - Pop_data_all$`2020`)/Pop_data_all$`2020`
Error2025 = abs(Pop_model$X2025 - Pop_data_all$`2025`)/Pop_data_all$`2025`
Error2030 = abs(Pop_model$X2030 - Pop_data_all$`2030`)/Pop_data_all$`2030`
Pop_model$Error2015 = Error2015
Pop_model$Error2020 = Error2020
Pop_model$Error2025 = Error2025
Pop_model$Error2030 = Error2030
Pop_model = mutate(Pop_model, mean_Error = (Error2015 + Error2020 + Error2025 + Error2030)/4)
summary(Pop_model$mean_Error)
```

```{r}
# PLots of errors
# Drop the NAs
Pop_model_err = drop_na(Pop_model)
ggplot(Pop_model_err, aes(x = "", y = mean_Error)) + 
  geom_boxplot() + 
  ylab('Error rate')
```

```{r}
# Using loess model
# Error compared with the given estimates
Error2015 = abs(Pop_model3$X2015 - Pop_data_all$`2015`)/Pop_data_all$`2015`
Error2020 = abs(Pop_model3$X2020 - Pop_data_all$`2020`)/Pop_data_all$`2020`
Error2025 = abs(Pop_model3$X2025 - Pop_data_all$`2025`)/Pop_data_all$`2025`
Error2030 = abs(Pop_model3$X2030 - Pop_data_all$`2030`)/Pop_data_all$`2030`
Pop_model3$Error2015 = Error2015
Pop_model3$Error2020 = Error2020
Pop_model3$Error2025 = Error2025
Pop_model3$Error2030 = Error2030
Pop_model3 = mutate(Pop_model3, mean_Error = (Error2015 + Error2020 + Error2025 + Error2030)/4)

# PLots of errors
# Drop the NAs
Pop_model_err3 = drop_na(Pop_model3)
hist(Pop_model_err3$mean_Error)
```



#### 3.4 Analysis of predictions 
```{r}
# Top 200 cities in 1950 and 2030
worldmap2 = getMap(resolution = "li")
top200_list_2030 = Pop_model_err[order(-Pop_model_err[[22]]),][1:200,]
top200_list_1950 = Pop_model_err[order(-Pop_model_err[[5]]),][1:200,]
plot(worldmap2)
points(top200_list_1950$Longitude, top200_list_1950$Latitude, col = "steelblue", pch = 20, cex = 1)
points(top200_list_2030$Longitude, top200_list_2030$Latitude, col = "brown1", pch = 20, cex = 0.5)
```

```{r}
# Countries in 1950 
country_table = as.data.frame(table(as.factor(top200_list_1950$Country.or.area)))
country_table = filter(country_table, Freq > 2)
country_table = country_table[order(country_table$Freq),]
# barplot(country_table$Freq, 
#        names.arg = country_table$Var1, main = 'Countries',
#        col = 'lightblue',
#        horiz = TRUE, cex.names = 0.5, las = 1,
#        xlim = c(0,400))
#axis(1, at = seq(0, 400, 25))

ggplot(data = country_table, aes(x = Var1, y = Freq)) +
  geom_bar(stat = 'identity', fill="lightblue")+
  geom_text(aes(label = Freq), vjust=1, color="black", size=3)+
  ylab("Frequency")+
  xlab("Country") +
  ggtitle(paste("Most resided countries"))+
  scale_x_discrete(limits = country_table$Var1)+
  theme_minimal() + 
  coord_flip()
```

```{r}
# Countries in 2030
country_table = as.data.frame(table(as.factor(top200_list_2030$Country.or.area)))
country_table = filter(country_table, Freq > 2)
country_table = country_table[order(country_table$Freq),]
# barplot(country_table$Freq, 
#        names.arg = country_table$Var1, main = 'Countries',
#        col = 'lightblue',
#        horiz = TRUE, cex.names = 0.5, las = 1,
#        xlim = c(0,400))
#axis(1, at = seq(0, 400, 25))

ggplot(data = country_table, aes(x = Var1, y = Freq)) +
  geom_bar(stat = 'identity', fill="lightblue")+
  geom_text(aes(label = Freq), vjust=1, color="black", size=3)+
  ylab("Frequency")+
  xlab("Country") +
  ggtitle(paste("Most resided countries"))+
  scale_x_discrete(limits = country_table$Var1)+
  theme_minimal() + 
  coord_flip()
```



```{r}
# Clustering of top 200 cities
k.max = 12
# tot.withinss = Total within-cluster sum of square
# iter.max = the maximum number of iterations allowed
# nstart = if centers is a number, how many random sets should be chosen.

# 1950
wss1 = sapply(1:k.max,function(k){kmeans(top200_list_1950[,3:4], k, nstart = 20, iter.max = 20)$tot.withinss})
plot(1:k.max,wss1, type= "b", xlab = "Number of clusters(k)", ylab = "Within cluster sum of squares")
icluster1 = kmeans(top200_list_1950[,3:4], 6, nstart = 7, iter.max = 20)

# 2030
wss2 = sapply(1:k.max,function(k){kmeans(top200_list_2030[,3:4], k, nstart = 20, iter.max = 20)$tot.withinss})
plot(1:k.max,wss2, type= "b", xlab = "Number of clusters(k)", ylab = "Within cluster sum of squares")
icluster2 = kmeans(top200_list_2030[,3:4], 6, nstart = 7, iter.max = 20)

# plot 
plot(worldmap2)
centers1 = as.data.frame(icluster1$centers)
centers2 = as.data.frame(icluster2$centers)
points(top200_list_1950$Longitude, top200_list_1950$Latitude, col = "lightblue", pch = 20, cex = 1)
points(top200_list_2030$Longitude, top200_list_2030$Latitude, col = "coral1", pch = 20, cex = 0.6)
points(centers1$Longitude, centers1$Latitude, col = "steelblue", pch = 20, cex = 3)
points(centers2$Longitude, centers2$Latitude, col = "brown1", pch = 20, cex = 3)
```

```{r}
# Top 20 cities in 1950 
year = "1950"
temp_data = Pop_model_err
temp_data = temp_data[order(-temp_data$X1950),] 
temp_20 = temp_data[1:20,]
temp_20 = temp_20[order(temp_20$X1950),] # For barplot in descending order
barplot(temp_20$X1950, 
        main=paste("Population of top 20 resided cities in", year), 
        names.arg = temp_20$Urban.Agglomeration, 
        horiz = TRUE, cex.names = 0.5, las = 2, col = 'lightblue',
        cex.axis = 0.8,
        xlab = 'Popolation (in thousands)',
        xlim = c(0, max(temp_20$X1950))) 
```

```{r}
# Top 20 cities in 2030
year = "2030"
temp_data = Pop_model_err
temp_data = temp_data[order(-temp_data$X2030),] 
temp_20 = temp_data[1:20,]
temp_20 = temp_20[order(temp_20$X2030),] # For barplot in descending order

ggplot(data = temp_20, aes(x = Urban.Agglomeration, y = X2030)) +
  geom_bar(stat = 'identity', fill="lightblue")+
  geom_text(aes(label = round(X2030)), vjust=1, color="black", size=3)+
  ylab("Population (in thousands)")+
  ylim(c(0,60000)) +
  ggtitle(paste("Top 20 resided cities in ", year))+
  scale_x_discrete(limits = temp_20$Urban.Agglomeration)+
  theme_minimal() + 
  coord_flip()
```


#### 3.5 Developing stages and classification
If the increasing rate has been decreased for more than three temporals, it will be classified as stage 3, if more than 1, it is in stage 2, otherwise it is stage 1. Then a classification is performed to identify the developing stage without access to all the data. 
```{r}
# Increasing rate for each city
# rate_all2 is the 2nd derivatives of each city
rate_all = data.frame(Pop_data_all[c(2,4,6:7,9:20)])
rate_all2 = rate_all
for(i in c(8:19)){
  j = i - 3 
  rate_all[,j] = (Pop_data_all[,i + 1] - Pop_data_all[,i])/(Pop_data_all[,i]+1)
}

for(i in c(6:15)){
    rate_all2[,i] = rate_all[,i] - rate_all[,i-1]
}
rate_all2$X1955 = NULL
rate_all2$X2010 = NULL
stage_vec = rep(0, nrow(Pop_data_all))
rate_all2$Stage = stage_vec 

for(i in c(1:nrow(Pop_data_all))){
  if(sum(rate_all2[i,5:14] < 0) > 1 & sum(rate_all2[i,5:14] < 0) < 4){
    rate_all2$Stage[i] = 2
  }else if(sum(rate_all2[i,5:14] < 0) > 3){
   rate_all2$Stage[i] = 3
  }else{
    rate_all2$Stage[i] = 1
  }
}
```

```{r}
table(rate_all2$Stage)
stage_df = as.data.frame(table(rate_all2$Stage))
bar = barplot(table(rate_all2$Stage), 
    main="Pie Chart of Development Stages",
    col = 'lightblue',
    xlab = 'Stage', 
    ylab = 'Frequency')
text(bar, stage_df$Var1 , paste(stage_df$Freq) ,cex=1, pos = 3) 
```

```{r}
# Training and testing set
# set.seed(0)
stage_all = data.frame(Pop_data_all[c(2,4,6:8)], rate_all2[15])
sample_num = sample(1692, 1500)
stage_train = stage_all[sample_num, ]
stage_test = stage_all[-sample_num, ]
```

```{r}
# Classification using the location and the population at 1950
# Using KNN
stage_knn = knn(train = stage_train[3:4], test = stage_test[3:4], cl = stage_train$Stage, k = round(sqrt(1500)))
table(stage_knn)
stage_knn_test = stage_test

stage_knn_test$prediction = stage_knn
stage_knn_test$pred_true = as.numeric(stage_knn_test$prediction == stage_knn_test$Stage)
paste('The accuracy rate of the prediction of KNN is', sum(stage_knn_test$pred_true)/nrow(stage_knn_test))
```

```{r}
table(stage_test$Stage, stage_knn_test$prediction)
```

```{r}
# Locations of the cities in different dtages
worldmap = getMap(resolution = "li")
plot(worldmap)
stage1 = stage_all[which(stage_all$Stage == 1),]
stage2 = stage_all[which(stage_all$Stage == 2),]
stage3 = stage_all[which(stage_all$Stage == 3),]
points(stage1$Longitude, stage1$Latitude, col = "brown1", pch = 20, cex = 0.3)
points(stage2$Longitude, stage2$Latitude, col = "lightgreen", pch = 20, cex = 0.3)
points(stage3$Longitude, stage3$Latitude, col = "steelblue", pch = 20, cex = 0.3)
```



```{r}
# K-Means
stage_kmeans_all = stage_all
icluster_stage = kmeans(stage_all[,3:4], 3, nstart = 7, iter.max = 20)
stage_kmeans_all$prediction = icluster_stage$cluster
stage_kmeans_all$pred_true = as.numeric(stage_kmeans_all$prediction == stage_kmeans_all$Stage)
paste('The accuracy rate of the prediction of KNN is', sum(stage_kmeans_all$pred_true)/nrow(stage_kmeans_all))
table(stage_kmeans_all$prediction, stage_kmeans_all$Stage)
```

